---
title: "Revisioning GWP of wetlands"
author: "Samuel D Chamberlain^1^"
date: '`r Sys.Date()`'
output:
  word_document: default
  pdf_document: default
  html_document: default
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
- \usepackage{hyperref}
fontsize: 12pt
---

^1^Department of Environmental Science, Policy, and Management, University of California, Berkeley, California, USA

```{r setup, include=FALSE}
#setwd to main package directory
knitr::opts_knit$set(root.dir = normalizePath("../"))
options(digits=2)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages, message=FALSE}
# packages used
library(knitr)
library(tidyverse)
library(gridExtra)
library(scales)
library(zoo)
library(lubridate)

# GWP* functions
source("R/co2e_model.R")
source("R/gwpfe_model.R")
source("R/emission_savings.R")
```

# Abstract

# 1. Introduction

# 2. Methods

# 3. Results
```{r aggregate_fluxes}
source("R/alluvium_processing.R")
source("R/peat6_processing.R")
source("R/peat19_processing.R")

#remove drawdown disturbance years from analysis
peat19_sub <- subset(peat19_yearly, year > 2013)
peat6_sub <- subset(peat6_yearly, year < 2017)
```

```{r fig1, fig.cap = "Mean annual fluxes via GWP"}
# Process all sites to mean annual fluxes with confidence intervals
all_sites <- rbind(peat19_sub, peat6_sub); all_sites <- rbind(all_sites, alluvium_yearly) # Combine all dataframes

ma_fluxes <- all_sites %>%
  group_by(site) %>%
  summarise(mCH4 = mean(tCH4),
            sdCH4 = sd(tCH4),
            seCH4 = sd(tCH4)/sqrt(length(tCH4)),
            mNEE = mean(tNEE),
            sdNEE = sd(tNEE),
            seNEE = sd(tNEE)/sqrt(length(tNEE)))


ggplot(ma_fluxes, aes(x = mNEE, y = mCH4, color = site)) +
  geom_point() +
  geom_errorbar(aes(ymax = mCH4 + seCH4, ymin = mCH4 - seCH4), width = 10) +
  geom_errorbarh(aes(xmax = mNEE + seNEE, xmin = mNEE - seNEE), height = 2) +
  geom_abline(intercept = 0, slope = -(1/28)*(44/16)) +
  scale_x_continuous(limits = c(-600, 0)) +
  scale_y_continuous(limits = c(0, 60))
  


```

```{r fig2, fig.width=7, fig.height=7, dpi=600}
# Conventional GWP budgets
mature <- gwpfe_model(time = 200, co2 = -424, co2sd = 223, ch4 = 48, ch4sd = 6.9, type = "eq", runs = 1000)
peat <- gwpfe_model(time = 200, co2 = -192, co2sd = 199, ch4 = 53, ch4sd = 10.4, type = "eq", runs = 1000)
alluvium <- gwpfe_model(time = 200, co2 = -320, co2sd = 392, ch4 = 32, ch4sd = 13.7, type = "eq", runs = 1000)

# GWP* forcing equivalent budgets
matureFE <- gwpfe_model(time = 200, co2 = -424, co2sd = 78.7, ch4 = 48, ch4sd = 6.9, type = "fe", runs = 1000)
peatFE <- gwpfe_model(time = 200, co2 = -192, co2sd = 199, ch4 = 53, ch4sd = 10.4, type = "fe", runs = 1000)
alluviumFE <- gwpfe_model(time = 200, co2 = -320, co2sd = 392, ch4 = 32, ch4sd = 13.7, type = "fe", runs = 1000)

a <- ggplot(peat) +
  geom_hline(yintercept = 0) +
  geom_ribbon(aes(x = time, ymin = (ch4_cum - ch4_sd)/1000, ymax = (ch4_cum + ch4_sd)/1000), fill = "salmon 1", alpha = 0.5) +
  geom_line(aes(x = time, y = ch4_cum/1000), color = "red") +
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "lightgreen", alpha = 0.5) +
  geom_line(aes(x = time, y = co2_cum/1000), color = "darkgreen") +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey" , alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  scale_y_continuous(limits = c(-250, 250)) +
  ylab("") + xlab("") + ggtitle("Peat wetland") +
  theme_bw()

b <- ggplot(peatFE) +
  geom_hline(yintercept = 0) +
  geom_ribbon(aes(x = time, ymin = (ch4_cum - ch4_sd)/1000, ymax = (ch4_cum + ch4_sd)/1000), fill = "salmon 1", alpha = 0.5) +
  geom_line(aes(x = time, y = ch4_cum/1000), color = "red") +
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "lightgreen", alpha = 0.5) +
  geom_line(aes(x = time, y = co2_cum/1000), color = "darkgreen") +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey" , alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  scale_y_continuous(limits = c(-250, 250)) +
  ylab("") + xlab("") + ggtitle("Peat wetland") +
  theme_bw()

c <- ggplot(mature) +
  geom_hline(yintercept = 0) +
  geom_ribbon(aes(x = time, ymin = (ch4_cum - ch4_sd)/1000, ymax = (ch4_cum + ch4_sd)/1000), fill = "salmon 1", alpha = 0.5) +
  geom_line(aes(x = time, y = ch4_cum/1000), color = "red") +
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "lightgreen", alpha = 0.5) +
  geom_line(aes(x = time, y = co2_cum/1000), color = "darkgreen") +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey" , alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  scale_y_continuous(limits = c(-250, 250)) +
  ylab("GWP CO2eq. emissions (kg m-2)") + xlab("") + ggtitle("Mature wetland") +
  theme_bw()

d <- ggplot(matureFE) +
  geom_hline(yintercept = 0) +
  geom_ribbon(aes(x = time, ymin = (ch4_cum - ch4_sd)/1000, ymax = (ch4_cum + ch4_sd)/1000), fill = "salmon 1", alpha = 0.5) +
  geom_line(aes(x = time, y = ch4_cum/1000), color = "red") +
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "lightgreen", alpha = 0.5) +
  geom_line(aes(x = time, y = co2_cum/1000), color = "darkgreen") +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey" , alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  scale_y_continuous(limits = c(-250, 250)) +
  ylab("GWP* CO2eq. emissions (kg m-2)") + xlab("") + ggtitle("Mature wetland") +
  theme_bw()

e <- ggplot(alluvium) +
  geom_hline(yintercept = 0) +
  geom_ribbon(aes(x = time, ymin = (ch4_cum - ch4_sd)/1000, ymax = (ch4_cum + ch4_sd)/1000), fill = "salmon 1", alpha = 0.5) +
  geom_line(aes(x = time, y = ch4_cum/1000), color = "red") +
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "lightgreen", alpha = 0.5) +
  geom_line(aes(x = time, y = co2_cum/1000), color = "darkgreen") +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey" , alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  scale_y_continuous(limits = c(-250, 250)) +
  ylab("") + xlab("Time since restoration (years)") + ggtitle("Alluvium wetland") +
  theme_bw()

f <- ggplot(alluviumFE) +
  geom_hline(yintercept = 0) +
  geom_ribbon(aes(x = time, ymin = (ch4_cum - ch4_sd)/1000, ymax = (ch4_cum + ch4_sd)/1000), fill = "salmon 1", alpha = 0.5) +
  geom_line(aes(x = time, y = ch4_cum/1000), color = "red") +
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "lightgreen", alpha = 0.5) +
  geom_line(aes(x = time, y = co2_cum/1000), color = "darkgreen") +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey" , alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  scale_y_continuous(limits = c(-250, 250)) +
  ylab("") + xlab("Time since restoration (years)") + ggtitle("Alluvium wetland") +
  theme_bw()

grid.arrange(a, b, c, d, e, f, ncol = 2)

```

```{r fig3, fig.height= 7, fig.width=4, dpi = 600}

corn_savings <- emission_savings(time = 150, baseline_co2 = 958.8, baseline_co2sd = 657.3, co2 = -324, co2sd = 78.7, ch4 = 48-1.7, ch4sd = 6.9, type = "fe", runs = 1000)
pasture_savings <- emission_savings(time = 150, baseline_co2 = 305.7, baseline_co2sd = 78.7, co2 = -324, co2sd = 78.7, ch4 = 48-9.1, ch4sd = 6.9, type = "fe", runs = 1000)
alfalfa_savings <- emission_savings(time = 150, baseline_co2 = 411.02, baseline_co2sd = 236.4, co2 = -324, co2sd = 78.7, ch4 = 48-0.5, ch4sd = 6.9, type = "fe", runs = 1000)

#identify the point at which wetland becomes GHG negative
corn_index <- which(corn_savings$net_mean < 0)[1]
pasture_index <- which(pasture_savings$net_mean < 0)[1]
alfalfa_index <- which(alfalfa_savings$net_mean < 0)[1]

a <- ggplot(corn_savings) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = corn_index, linetype = "dashed", size = 0.5) +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey", alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  geom_ribbon(aes(x = time, ymin = (co2base_cum - co2base_sd)/1000, ymax = (co2base_cum + co2base_sd)/1000), fill = "yellow2" , alpha = 0.5) +
  geom_line(data = corn_savings, aes(x = time, y = co2base_cum/1000), color = "yellow4") +  
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "green", alpha = 0.5) +
  geom_line(data = corn_savings, aes(x = time, y = co2_cum/1000), color = "dark green") +  
  ylab("CO2eq. budget (kg m-2)") + xlab("Time since restoration (years)") +
  #annotate("text", x = 50, y = 185, label = "Corn: 111 kg savings @ 100yr") +
  scale_y_continuous(limits = c(-200, 200)) +
  theme_bw()

b <- ggplot(pasture_savings) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = pasture_index, linetype = "dashed", size = 0.5) +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey" , alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  geom_ribbon(aes(x = time, ymin = (co2base_cum - co2base_sd)/1000, ymax = (co2base_cum + co2base_sd)/1000), fill = "orange" , alpha = 0.5) +
  geom_line(aes(x = time, y = co2base_cum/1000), color = "orchid") +
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "green", alpha = 0.5) +
  geom_line(data = corn_savings, aes(x = time, y = co2_cum/1000), color = "dark green") +  
  ylab("CO2eq. budget (kg m-2)") + xlab("Time since restoration (years)") +
  ylab("CO2eq. budget (kg m-2)") + xlab("Time since restoration (years)") +
  #annotate("text", x = 50, y = 185, label = "Pasture: 52 kg savings @ 100yr") +
  scale_y_continuous(limits = c(-200, 200)) +
  theme_bw()

c <- ggplot(alfalfa_savings) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = alfalfa_index, linetype = "dashed", size = 0.5) +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey" , alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  geom_ribbon(aes(x = time, ymin = (co2base_cum - co2base_sd)/1000, ymax = (co2base_cum + co2base_sd)/1000), fill = "coral1" , alpha = 0.5) +
  geom_line(aes(x = time, y = co2base_cum/1000), color = "coral") + 
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "green", alpha = 0.5) +
  geom_line(data = corn_savings, aes(x = time, y = co2_cum/1000), color = "dark green") +  
  ylab("CO2eq. budget (kg m-2)") + xlab("Time since restoration (years)") +
  #annotate("text", x = 50, y = 185, label = "Alfalfa: 34 kg savings @ 100yr") +
  scale_y_continuous(limits = c(-200, 200)) +
  theme_bw()

grid.arrange(a, b, c, ncol = 1)
```

```{r fig4}
wetland <- co2e_model(time1 = 0, time2 = 150, pre_co2 = 305.7, pre_co2sd = 78.7, pre_ch4 = 9.1, pre_ch4sd = 4.0, 
                    pst_co2 = 449.4, pst_co2sd = 206.0, pst_ch4 = 44.9, pst_ch4sd = 8.6, runs = 1000)

ggplot(wetland) +
  geom_hline(yintercept = 0) +
  geom_ribbon(aes(x = time, ymin = (ch4_cum - ch4_sd)/1000, ymax = (ch4_cum + ch4_sd)/1000), fill = "salmon 1", alpha = 0.5) +
  geom_line(aes(x = time, y = ch4_cum/1000), color = "red") +
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "lightgreen", alpha = 0.5) +
  geom_line(aes(x = time, y = co2_cum/1000), color = "darkgreen") +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey" , alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  scale_y_continuous(limits = c(-250, 250)) +
  ylab("Cumulative CO2eq. emissions (kg m-2)") + xlab("Time since restoration (years)") + ggtitle("West") +
  theme_bw()
```

# 4. Discussion

# Acknowledgements
This research was supported in part by the U.S. Department of Energy’s Office of Science and its funding of Ameriflux core sites (Ameriflux contract 7079856), and the California Division of Fish and Wildlife, through a contract of the California Department of Water Resources (Award 4600011240).

# References

