---
title: Revisioning GWP of wetlands
author: Samuel D Chamberlain^1^
date: '`r Sys.Date()`'
output:
  word_document: default
  pdf_document: default
  html_document: default
fontsize: 12pt
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
- \usepackage{hyperref}
link-citations: yes
csl: institute-of-physics-harvard.csl
bibliography: library.bib
---

^1^Department of Environmental Science, Policy, and Management, University of California, Berkeley, California, USA

```{r setup, include=FALSE}
#setwd to main package directory
knitr::opts_knit$set(root.dir = normalizePath("../"))
options(digits=2)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages, message=FALSE}
# packages used
library(knitr)
library(tidyverse)
library(gridExtra)
library(scales)

# GWP* functions
source("R/co2e_model.R")
```

# Abstract

# 1. Introduction

# 2. Methods

# 3. Results

```{r fig1}
wetland <- co2e_model(time1 = 0, time2 = 150, pre_co2 = 305.7, pre_co2sd = 78.7, pre_ch4 = 9.1, pre_ch4sd = 4.0, 
                    pst_co2 = 449.4, pst_co2sd = 206.0, pst_ch4 = 44.9, pst_ch4sd = 8.6, runs = 1000)

ggplot(wetland) +
  geom_hline(yintercept = 0) +
  geom_ribbon(aes(x = time, ymin = (ch4_cum - ch4_sd)/1000, ymax = (ch4_cum + ch4_sd)/1000), fill = "salmon 1", alpha = 0.5) +
  geom_line(aes(x = time, y = ch4_cum/1000), color = "red") +
  geom_ribbon(aes(x = time, ymin = (co2_cum - co2_sd)/1000, ymax = (co2_cum + co2_sd)/1000), fill = "lightgreen", alpha = 0.5) +
  geom_line(aes(x = time, y = co2_cum/1000), color = "darkgreen") +
  geom_ribbon(aes(x = time, ymin = (net_mean - net_sd)/1000, ymax = (net_mean + net_sd)/1000), fill = "light grey" , alpha = 0.5) +
  geom_line(aes(x = time, y = net_mean/1000)) +
  scale_y_continuous(limits = c(-250, 250)) +
  ylab("Cumulative CO2eq. emissions (kg m-2)") + xlab("Time since restoration (years)") + ggtitle("Pasture") +
  theme_bw()
```

# 4. Discussion

# Acknowledgements
This research was supported in part by the U.S. Department of Energy’s Office of Science and its funding of Ameriflux core sites (Ameriflux contract 7079856), and the California Division of Fish and Wildlife, through a contract of the California Department of Water Resources (Award 4600011240).

# References

